---
title: 'Regular Season vs Playoffs: Shooting Efficiency Analysis'
author: "STAT 107 Group 20"
date: "2025-11-08"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)

# Load required packages listed in 00_requirements.R
source("00_requirements.R")
```

# Abstract
- This report examines whether NBA players’ shooting efficiency changes between the regular season and the playoffs. Using player level data from the Oklahoma City Thunder, we focus on field goal percentage and three point percentage. We rely on a dedicated cleaning file to standardize variables and on a visualization file to produce distribution, paired comparison, and correlation plots. Preliminary visuals suggest modest declines in playoff efficiency. In the final phase we will run paired tests and report effect sizes with confidence intervals.

# Introduction
- Playoffs often feature stronger defenses and higher pressure. These conditions may affect shooting outcomes. We ask whether field goal percentage and three point percentage differ between the two stages for the same players and whether regular season performance correlates with playoff performance.

# Data and Cleaning
- The datasets are Regular Season.csv and Playoffs.csv. Cleaning and variable standardization are performed in the separate R Markdown file 11_DataCleaning.Rmd.
- Step 1: Load the datasets. First, both the regular-season and playoff datasets were imported into the workspace. After loading them, we checked the structure of each dataset to understand what variables they included, how the data were formatted, and whether there were any immediate issues. This gave us a baseline understanding of the raw data before making any changes.
- Step 2: Keep only the relevant columns. Next, we reduced each dataset to just the variables needed for the analysis. This step removed extra or unnecessary columns so that the remaining datasets contained only the player identity, team, playing time, scoring totals, and shooting statistics we planned to work with.
- Step 3: Rename columns for consistency. Because the column names in the two datasets were not fully consistent, all selected variables were renamed to follow the same naming style. This standardization ensured that both datasets used identical labels, which is important for combining them later and for avoiding confusion during analysis.
- Step 4: Combine the datasets. Finally, the cleaned regular-season and playoff datasets were merged into a single dataset. A new variable was added to label each observation as either “Regular” or “Playoffs,” making it possible to compare player performance across the two contexts within one unified dataset.

# Visualization
- We include the visualization plots that visualize the distribution of shooting efficiency, paired boxplots for player-level performance comparison, and correlation scatter plots between regular and playoff FG%. The plotting functions are defined in 02_funct_Plots.R and are used inside the visualization script.
```{r child="31_DataVisualization.Rmd", echo=FALSE, warning=FALSE, message=FALSE}

```

# Summary Statistics
- The table below summarizes group level statistics (mean and standard deviation) for FG% and 3P% across both stages. These statistics are based on the combined dataset that was created and prepared in the 11_DataCleaning.Rmd file.
```{r summary-stats, message=FALSE, warning=FALSE}

# Compute and summarize overall FG% and 3P% by stage
desc <- combined %>%
  dplyr::group_by(type) %>% # Group by stage (Regular or Playoffs)
  dplyr::summarise(
    players_n = dplyr::n(),
    FG_mean = mean(FG_percent, na.rm = TRUE),
    FG_sd = sd(FG_percent, na.rm = TRUE),
    TP_mean = mean(TP_percent, na.rm = TRUE),
    TP_sd = sd(TP_percent, na.rm = TRUE),
    .groups = "drop"
  )

# Display results as a formatted table in the report
knitr::kable(desc, caption = "Summary statistics for FG% and 3P% by stage")
```




