---
title: 'Regular Season vs Playoffs: Shooting Efficiency Analysis'
author: "STAT 107 Group 20"
date: "2025-11-08"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)

# Load required packages listed in 00_requirements.R
source("00_requirements.R")
```

# Abstract
- This report examines whether NBA players’ shooting efficiency changes between the regular season and the playoffs. Using player level data from the Oklahoma City Thunder, we focus on field goal percentage and three point percentage. We rely on a dedicated cleaning file to standardize variables and on a visualization file to produce distribution, paired comparison, and correlation plots. Preliminary visuals suggest modest declines in playoff efficiency. In the final phase we will run paired tests and report effect sizes with confidence intervals.

# Introduction
- Playoffs often feature stronger defenses and higher pressure. These conditions may affect shooting outcomes. We ask whether field goal percentage and three point percentage differ between the two stages for the same players and whether regular season performance correlates with playoff performance.

# Data and Cleaning
- The datasets are Regular Season.csv and Playoffs.csv. Cleaning and variable standardization are performed in the separate R Markdown file 11_DataCleaning.Rmd.
```{r child="11_DataCleaning.Rmd", echo=FALSE, warning=FALSE, message=FALSE}

```

# Visualization
- We include the visualization plots that visualize the distribution of shooting efficiency, paired boxplots for player-level performance comparison, and correlation scatter plots between regular and playoff FG%. The plotting functions are defined in 02_funct_Plots.R and are used inside the visualization script.
```{r child="31_DataVisualization.Rmd", echo=FALSE, warning=FALSE, message=FALSE}

```

# Summary Statistics
- The table below summarizes group level statistics (mean and standard deviation) for FG% and 3P% across both stages. These statistics are based on the combined dataset that was created and prepared in the 11_DataCleaning.Rmd file.
```{r summary-stats, message=FALSE, warning=FALSE}

# Compute and summarize overall FG% and 3P% by stage
desc <- combined %>%
  dplyr::group_by(type) %>% # Group by stage (Regular or Playoffs)
  dplyr::summarise(
    players_n = dplyr::n(),
    FG_mean = mean(FG_percent, na.rm = TRUE),
    FG_sd = sd(FG_percent, na.rm = TRUE),
    TP_mean = mean(TP_percent, na.rm = TRUE),
    TP_sd = sd(TP_percent, na.rm = TRUE),
    .groups = "drop"
  )

# Display results as a formatted table in the report
knitr::kable(desc, caption = "Summary statistics for FG% and 3P% by stage")
```



